---
title: 'R project: SLR and Bootstrap'
author: "Lince Rumainum"
date: "13 December 2020"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction:
What problem do you wish to solve with `myslr()`? -- 
Function that creates estimates for the betas in a  slr.
The function will take an x vector and a y vector.
Scatter Plot shows the relation of y to x value and create fitted line 
Histograms shows x and y data
Output csv file shows comparison of predictions using built-in function and slr calculation results
Commandline output shows input data, plots, and betas values from both built-in function and slr calculation

# Data
We will all apply our function to the mtcars data set and make a linear model `mpg ~ wt` - but remember the function will be completely general to SLR. 

Describe the data

Describe the nature and type of the variables.
The mtcars data frame consist of 11 numeric variables with 32 data observations

The variables are:
[ 1]	mpg	 -- Miles/(US) gallon
[ 2]	cyl	 -- Number of cylinders
[ 3]	disp -- Displacement (cu.in.)
[ 4]	hp	 -- Gross horsepower
[ 5]	drat -- Rear axle ratio
[ 6]	wt	 -- Weight (1000 lbs)
[ 7]	qsec -- 1/4 mile time
[ 8]	vs	 -- Engine (0 = V-shaped, 1 = straight)
[ 9]	am	 -- Transmission (0 = automatic, 1 = manual)
[10]	gear --	Number of forward gears
[11]	carb --	Number of carburetors

Plot the data
First plot (default):
It shows that the heavier the vehicle, the less efficient (mpg) it would be. It also shows the linear model from the observation points with its range of variance.
Second plot:
It shows the histograms of mpg data and separate them between engine type (0 = V-shaped, 1 = straight)
It shows that V-shaped engine most likely to have lower mpg than straight engine
Most of the V-shaped engine has approximately 15 mpg while the lowest mpg for straight engine is approximately 17.5 mpg

```{r cars}
data(mtcars) #load data
head(mtcars) #look up first few rows of data 

library(ggplot2) #load ggplot2
#plot mtcars data mpg vs wt with cyl as fill
#plot points and smooth (abline and variance shadowed) from data
g = ggplot(mtcars, aes(x = wt, y=mpg, fill = cyl)) + geom_point() + geom_smooth(method="lm", formula=y~x)
g

#create histogram for mpg data 
h = ggplot(mtcars, aes(x=mpg, fill=..density..))
#create density on y
h = h + geom_histogram(aes(y=..density..))
#color red for high density and green for low density
h = h + scale_fill_gradient("density", low="green", high="red")
#have facet grid for the engine type
#0 - V-shaped and 1 - straight
h = h + facet_grid(vs~.)
#create a dark theme background for the plots
h = h + theme_dark()
#plot
h
```

# Theory used
Give the mathematical formulae in Latex.
Interpret the meaning of the symbols.

$\hat{Y}_i$ = $\hat{\beta}_0$ + $\hat{\beta}_1 X_i$


$\hat{Y}_i$ - the predicted variable; dependent of given value of x
$\hat{\beta}_0$ - the y-intercept of the regression line
$\hat{\beta}_1$ - the x-interval coefficient / slope

Predictions of $\hat{\beta}_0$ and $\hat{\beta}_1$ can be calculated from the linear regression equation below:

$\hat{\beta}_0$ = [ ($\sum$y)($\sum$x$^2$) - ($\sum$x)($\sum$xy) ] / [ n($\sum$x$^2$) - $(\sum$ x)$^2$ ]

$\hat{\beta}_1$ = [ n($\sum$xy) - ($\sum$x)($\sum$y) ] / [ n($\sum$x$^2$) - $(\sum$ x)$^2$ ]

# Application of SLR to the mtcars data set
Use R to analysze the data
`y.lm=lm(y~x,data=mtcars)`
Check assumptions
`normcheck()`
`plot(y.lm, which =1)`

Interpret the summary output.
`summary(y.lm)`
`ciReg(y.lm)`
What are the point and interval estimates?
From summary:
The intercept point is 10.2779 with interval estimates of 4.9193
The R-squared of this simple linear model:
Multiple R-squared: 0.9235
Adjusted R-squared: 0.9176 
Since the R-squared value is close to 1, this linear model is considered to be a good model

With Confidence Intervals For Regression Models results:
              95 % C.I.lower    95 % C.I.upper
(Intercept)        7.20960          13.34625
x                  4.07085           5.76781

# Now create functions:

## Decide on the output you want

### Plots
  
  1) Plot the data with the fitted line 
  2) The plot can be made to appear when the function is used.
  3) The plot is saved to the working directory automatically.

### Files created

Files created for x-value and two prediction values. One variable contain predictions from the betas results of the built-in linear model function and another variable contain predictions from the betas results of the betas calculation based on the linear regression equations 

### Command line

Shows the input x and y, beta estimates and R-squared values from summary and the Confidence Intervals

# myslr

```{r myslr}
myslr = function(x,y,df)
{
  #load library
  library(s20x)
  library(ggplot2)
  library(dplyr)
  
  #create the linear model with built-in function
  y.lm = lm(y~x)
  
  #Create scatter plot of y vs x
  s = ggplot(df, aes(x=x, y = y))
  #plot points
  s = s + geom_point() 
  #plot linear model absolute line 
  s = s + geom_smooth(method="lm", formula=y~x) 
  #add title
  s = s + ggtitle("Scatter Plot")
  s = s + theme_gray()
  #save scatter plot 
  ggsave(path = getwd(), filename = "ScatterPlot.png",  dpi = 1000)
  
  #create histogram plot 
  #histogram of x
  #create histogram for x data 
  hx = ggplot(df, aes(x=x, fill=..density..))
  #create density on y-axis
  hx = hx + geom_histogram(aes(y=..density..))
  #color red for high density and green for low density
  hx = hx + scale_fill_gradient("density", low="green", high="red")
  #create a dark theme background for the plots
  hx = hx + theme_dark()
  #save x histogram 
  ggsave(path = getwd(), filename = "x-Histogram.png",  dpi = 1000)
  
  #histogram of y
  #create histogram for y data 
  hy = ggplot(df, aes(x=y, fill=..density..))
  #create density on y-axis
  hy = hy + geom_histogram(aes(y=..density..))
  #color red for high density and green for low density
  hy = hy + scale_fill_gradient("density", low="green", high="red")
  #create a black/white theme background for the plots
  hy = hy + theme_bw()
  #save y histogram
  ggsave(path = getwd(), filename = "y-Histogram.png",  dpi = 1000)
  
  #get the confidence interval results 
  #and summary of the linear model
  ci.lm = ciReg(y.lm)
  summary.lm = summary(y.lm)
  
  #Calculate estimation of betas from linear regression equation

  #create variables
  sum_y = sum(y)    #sum of y
  sum_x2 = sum(x^2) #sum of x^2 
  sum_x = sum(x)    #sum of x
  sum_xy = sum(x*y) #sum of x*y
  
  n <- length(x) # get number of data
  
  #Calculate b0 (y-Intercept) and b1 (x-coefficient)
  b0 <- (sum_y*sum_x2 - sum_x*sum_xy) / (n*sum_x2 - (sum_x^2))
  b1 <- (n*sum_xy - sum_x * sum_y) / (n*sum_x2 - (sum_x^2))
  
  #create x_list with increment of 0.1 
  x_list <- as.data.frame(seq(min(x)-1,max(x)+1,0.1))
  
  #create y_prediction from betas results of built-in function
  y_pred <- as.data.frame(y.lm$coef[1] + y.lm$coef[2] * x_list)
  
  #create y_prediction from betas results of equation
  y_pred_calc <- as.data.frame(b0 + b1 * x_list)
  
  #unlist data and convert as numeric
  x_points <- as.numeric(unlist(x_list))
  y_built_in_func <- as.numeric(unlist(y_pred))
  y_calculation <- as.numeric(unlist(y_pred_calc))
  
  #create results data with x and  
  #two y's values (one betas from built-in lm and one betas from calculation)
  result_data <- tibble(x=x_points,y_lm=y_built_in_func,y_calc=y_calculation)
  
  path <- getwd()
  resultFile <- paste(path, "\\results.csv", sep = "")
  write.csv(result_data, resultFile, row.names = F)
  
  #command line list of:
  #the x and y inputs 
  #the beta estimates from :
  #the confidence interval and the summary of the linear model
  list(x=x, y=y,
       scatterPlot = s, 
       histogram_x = hx, 
       histogram_y = hy, 
       confindence_interval=ci.lm, 
       summary=summary.lm,
       yIntercept_from_calculation = b0,
       xCoefficient_from_calculation = b1)
}

```

# Now invoke your functiion

```{r invokeslr, eval=TRUE}
#load library
library(s20x)
library(dplyr)
#get x-data
x_dat = mtcars$wt
#get y-data
y_dat = mtcars$mpg
#create object for myslr function
obj1 = myslr(x=x_dat,y=y_dat,df=mtcars)
#view object data
obj1

```


# Bootstrap 

Make a function `myboot` that will create  bootstrap estimates from a sample.
The parameter estimated will be the population variance $\sigma^2$ of mpg for 4 cylinder cars. Though the function will be entirely general.

# Make the function

x is a sample vector alpha is the error rate, the `100(1-alpha)%`  confidence interval is `quantile(stat, c(alpha/2,1-alpha/2))`

## Plots 
histogram of the simulated statistic.
What statistic will you use to estimate $\sigma^2$

## Commandline

list of interesting objects like $\hat{\sigma^2}$

## File

write a file to the commandline that contains the simulation.

```{r myboot}
myboot = function(x, alpha)
{
  #load library
  library(s20x)
  library(dplyr)
  library(gridExtra)
  
  #get number of population
  n <- nrow(x)
  
  #initialize sample size, 100 x bigger than population since replacement is possible
  sample_size = n * 100
  #unlist x data
  x_vector = unlist(x)
  #create sample with replacement
  smple = sample(x_vector, sample_size, replace = TRUE)
  
  #create histogram of the density of population and sample
  
  #POPULATION
  #create data frame for input to be able to be plotted
  df_pop = as.data.frame(x)
  x_pop = unlist(df_pop[1])
  #create histogram with ggplot
  hd_pop = ggplot(df_pop, aes(x=x_pop, fill=..density..))
  #h_pop = ggplot(df_pop, aes(x=x_pop))
  #use the frequency on y-axis
  hd_pop = hd_pop + geom_histogram(aes(y=..density..))
  #h_pop = h_pop + geom_histogram(frequency=TRUE)
  #color red for high density and black for low density
  hd_pop = hd_pop + scale_fill_gradient("density", low="black", high="red")
  #create a gray theme background for the plots
  hd_pop = hd_pop + theme_gray()
  #plot
  hd_pop
  #save population's histogram
  ggsave(path = getwd(), filename = "pop-density-Histogram.png",  dpi = 1000)
  
  #SAMPLE
  #create sample data frame
  df_sample = as.data.frame(smple)
  #create x-samplefor histogram plot
  x_sample=unlist(df_sample[1])
  
  #create histogram with ggplot
  hd_sample = ggplot(df_sample, aes(x=x_sample, fill=..density..))
  #h_sample = ggplot(df_sample, aes(x=x_sample))
  #use the frequency on y-axis
  hd_sample = hd_sample + geom_histogram(aes(y=..density..))
  #h_sample = h_sample + geom_histogram(frequency=TRUE)
  #color red for high density and black for low density
  hd_sample = hd_sample + scale_fill_gradient("density", low="black", high="red")
  #create a gray theme background for the plots
  hd_sample = hd_sample + theme_dark()
  #plot
  hd_sample
  #save sample's histogram
  ggsave(path = getwd(), filename = "sample-density-Histogram.png",  dpi = 1000)
  
  #COMPARE HISTOGRAMS
  #create grid to put histogram of population and sample side-by-side
  hd_pop_sample = grid.arrange(hd_pop, hd_sample, ncol=2)
  #save stat's histogram
  ggsave(path = getwd(), filename = "pop-sample-density-Histogram.png", hd_pop_sample,  dpi = 1000)
  
  #create histogram of the frequency of population and sample
  
  #POPULATION
  #create histogram with ggplot
  hf_pop = ggplot(df_pop, aes(x=x_pop))
  #use the frequency on y-axis
  hf_pop = hf_pop + geom_histogram(frequency=TRUE)
  #create a gray theme background for the plots
  hf_pop = hf_pop + theme_gray()
  #plot
  hf_pop
  #save population's histogram
  ggsave(path = getwd(), filename = "pop-freq-Histogram.png",  dpi = 1000)
  
  #SAMPLE
  #create histogram with ggplot
  hf_sample = ggplot(df_sample, aes(x=x_sample))
  #use the frequency on y-axis
  hf_sample = hf_sample + geom_histogram(frequency=TRUE)
  #create a dark theme background for the plot
  hf_sample = hf_sample + theme_dark()
  #plot
  hf_sample
  #save stat's histogram
  ggsave(path = getwd(), filename = "sample-freq-Histogram.png",  dpi = 1000)
  
  #create grid to put histogram of population and sample side-by-side
  hf_pop_sample = grid.arrange(hf_pop, hf_sample, ncol=2)
  #save stat's histogram
  ggsave(path = getwd(), filename = "pop-sample-frequency-Histogram.png", hf_pop_sample,  dpi = 1000)
  
  #Calcuate statistics data below:
  
  #create object quantile result for the specified confidence interval
  quantile_result <- quantile(data.matrix(x_sample),c(alpha/2,1-alpha/2))
  
  #unlist x and convert as numeric for population variance calculation
  x_num = as.numeric(unlist(x[1]))
  #calculate population variance
  #using the population variance formula 
  pop_mean = mean(x_num)
  pop_var <- (sum((x_num-pop_mean)^2))/n
  #calculate sample variance
  #using the sample variance formula 
  #(same result as using var function from stat package --> var(x_stat))
  sample_mean = mean(x_sample)
  sample_var <- (sum((x_sample-sample_mean)^2))/(sample_size-1)
  
  #command line list of:
  #input data / population data
  #histogram of the density of population and sample data
  #sample variance of the sample data
  #population variance of the input data
  #histogram of the frequency of population and sample data
  #quantile result with specified confidence interval 
  list(input_data = x, 
       histogram_density_population_and_sample = hd_pop_sample,
       population_variance = pop_var, sample_variance = sample_var,
       histogram_frequency_population_and_sample = hf_pop_sample, 
       quantile = quantile_result)
}

```

# Invoke your function for the mtcars dataset

Must get x by using `subset()` or something else

```{r, eval=TRUE}
#load library
library(s20x)
library(dplyr)
#create a subset for mtcars's mpg data for 4 cylinders vehicles
sample_mpg_4cyl = mtcars %>% filter(cyl == 4) %>% select(mpg)
#create object using myboot function
obj = myboot(x= sample_mpg_4cyl, alpha = 0.05)
#view object's data
obj
```

# Make a <5 min recording of both functions running on the data sets with your explanation

You can use quicktime .mov for a mac or use BBFLASH free edition for PC -- make as a small file (keep as .fbr)

[https://www.flashbackrecorder.com/express/download/](https://www.flashbackrecorder.com/express/download/)
